>

### 2019-11-06 manuscript/manuscript.Rmd
Add appendix
Edits suggested by Kari

output: 
  rticles::elsevier_article:
    includes: 
      in_header: preamble.tex
csl: elsevier-harvard.csl
bibliography: library.bib
---

@@

The large majority of extant research on the topic of park access and public health
outcomes has been involved comparing metropolitan regions against each other. For example,
@West2012 used park data from
the Trust for Public Land’s 2010 "City Park Facts" and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine the relationship
between the density of park land and physical activity and obesity rates for 67
metropolitan statistical areas in the U.S. The findings in this case conformed to 
expectation; that is, the study found a significant positive association
between park density and physical activity and a significant negative
association between park density and obesity.
@Larson2016 similarly used self-reported scores on the Gallup-Healthways
Wellbeing Index to evaluate the relationship between physical health and park
quantity, quality, and accessibility in 44 U.S. cities. While the authors found
positive associations between wellbeing and park quality and accessibility,
these relationships were not statistically significant.
Conversely, @Richardson2012 examined the relationship between total urban
greenspace and mortality rates for selected maladies; though the authors did not
find a statistically significant relationship between the quantity of urban
greenspace and mortality caused by lung cancer, diabetes, heart disease or car
accidents, they did find that all-cause mortality was --- oddly --- *higher* in greener
cities. Finally, in a meta-analysis of 20 peer reviewed journal articles exploring the
relationship between parks and objectively measured physical activity,
@Bancroft2015 found that five studies reported a significant positive
association between the two, six studies produced mixed results, and nine
studies found no association at all.

@@

Variations of this isochrone-based framework are easily derived and relatively
common. 
ParkScore [@parkscore2019] is a sophisticated application of this approach where $D$ is a
carefully calculated 10-minute network-bound walk, paired with a demographic analysis of
areas within and without this threshold. In other applications $D$ might be more
variably defined, such as the presence or percent of green space within a
political or statistical boundary [@Mitchell2008; @Stark2014] or within a specific distance
threshold [@Kaczynski2014]. It is also possible to attenuate an isochrone with 
various factors; for instance, @Dias2019 considered road safety perception as an
extenuating factor in the effective distance to a park.

@@

In spite of its flexibility and basis in choice theory, utility-based
accessibility measures have not received much application in the accessibility
literature compared with distance-based or even gravity-based measures.
@Vale2016 present a descriptive classification of active accessibility techniques, and
explicitly dismiss utility-based metrics for incorporating randomness. The
accessibility formula presented in \ref{eq:utility} is the expectation of a
random utility process, and is not in and of itself random any more than the
gravity model is random.
Utility-based accessibility metrics are commonly used, however, in alternatives
analyses of transit infrastructure improvements [@DeJong2007]. A reason for this
is likely that a regional travel demand model is available to the analysts, thus
making calibrated and multimodal logsums readily available [@Geurs2010].

@@

To this dataset of parks and regularized green spaces, we add a polygons layer of
cemeteries in New York that are open to the public. Cemeteries are important
green spaces that can be used for many types of physical activity. These
operations leave us with `r nrow(parks)` distinct green spaces; for simplicity 
we will refer to both parks and cemeteries as "parks" going forward. 
For each park and 
we calculate the size of the park in acres. Using a Twitter application
programming interface (API) implemented with the Python package Tweepy,  we also
collected and stored tweets containing geotags of precise geographic coordinates
located within the boundaries of the parks; from this information,
we were able to segregate tweets that were generated within a park in September
2014. This geolocated twitter activity could provide an additional point of information
on the degree to which individual parks are actually used [@Wang2016].

@@

We calculate a set of utility-based accessibility statistics for each tract
in New York City. The most basic utility specification includes only the
the park size in acres and the distance from the population-weighted tract centroid
to the boundary of the park in miles,
\begin{equation}\label{eq:u_access}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij})
\end{equation}
The logarithmic transform allows for diminishing marginal utility of distance
and park size: A 1-mile increase to a trip matters more for a 1-mile trip than a
10-mile trip. @Macfarlane2019 estimated destination choice 
parameters for park trips in the Alameda County (Oakland), California using
this utility specification, obtaining values of $\lambda_s = 0.373$ and 
$\lambda_d = -1.76$. The ratio of these estimates implies people are willing to
travel roughly six times further to reach a park twice as large. In the absence 
of park trip destination choice coefficients in New York City, we adopt these
previous estimates.

@Kinnell2006 surveyed park users in New Jersey and applied a multinomial logit
model to determine which factors influenced its perceived utility as a park
trip destination. By transferring estimates of common variables between our
data and this survey-based model, we can create a second utility specification
as
\begin{equation}\label{eq:u_multi}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \lambda_t * trails + \\
  \lambda_{c} * courts + \lambda_p * playgrounds
\end{equation}
with $\lambda_t = 0.99$, $\lambda_c = 0.43$, and $\lambda_p = 0.26$. Kinnell et al.\
did not transform their size and distance estimates, so we retain the previously 
estimated and applied values. We assume that the other covariates in the 
model which we do not have available to us (e.g., boat launches, picnic areas) are
orthogonal to the included parameters and leaving them out will not affect the
values of the included coefficients.

The third utility specification includes the number of geolocated tweets
emanating from within the park in September of 2014 in addition to the
basic specification in Equation \ref{eq:u_access},
\begin{equation}\label{eq:u_tweets}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \lambda_{tw} * YJ(tweets)
\end{equation}
with $YJ(x)$ being a Yeo-Johnson transformation that implements a diminishing
marginal return similar to $log(x)$, but where $YJ(0) = 0$ [@Yeo2000]. In this
case we have no external information describing how destination choice to
parks may be affected by the twitter activity in the park, so we assert a 
value of $\lambda_{tw} = 0.1$

@@

As the observations are related to each other spatially, it is likely that this
process involves spatial spillovers of one kind or another. A complete treatment of
spatial econometrics is not warranted here; the reader is directed to
@LeSage2009 as well as @LeSage2014.
Suffice it to say that the presence of spatial data generating processes can
negatively affect econometric interpretation in a variety of ways. Each of these
process relies on spatial autoregression, where elements of some random
variable $x$ are spatially dependent on other elements, 
\begin{equation}\label{eq:autoregression}
 x = \rho W x + \varepsilon
\end{equation}
with the strength of dependence an estimated parameter $\rho$, the structure of 
spatial dependence governed by the asserted matrix $W$ (details of $W$ follow below),
and some random uncorrelated residual $\varepsilon$. A collection of models is
available to represent spatial autodependence in the model residuals, the 
independent variables, and the dependent variable.

@@

Table \ref{tab:pa-models} present the estimated spatial lag and accessibility
coefficients for three models using different specifications of the utility
equation, as well as the ten-minute walk buffer calculated by ParkScore. 
Note that the model called "Size and Distance" is precisely the model
labeled "SDEM" in Table \ref{tab:slx-sdem}. The controlling covariates are
suppressed for clarity but do not change meaningfully across the four specifications.
Complete model estimates are available from the authors. Each of the three
utility-based specifications is significant at the $\alpha = 0.05$ level or lower,
and is of the hypothesized directionality. That is, residents living in tracts 
with increased utility-based accessiblity to green spaces show a significantly
higher physical activity participation rate. Taking the rough mean of the three
estimated utility-based coefficient, moving from the least-accessible tract ($A_i = -3$) to the
most-accessible tract ($A_i = 3$) is expected to raise the physical activity rate
by `r 6 * .2` percentage points. With the understanding of the full 95\% confidence
interval, we could suggest that the true effect of excellent versus poor park access
is anywhere between `r 6 * .05` and `r 6 * .35` percentage points.

@@

We now consider the impacts of a model where the dependent variable is
the *obesity* rate in a tract, and physical activity becomes an independent covariate
alongside the controlling variables and the accessibility metrics.
Table \ref{tab:obesity-models} presents estimates of the effect of all three
utility-based accessibility measures as well as the 10-minute walk buffer. 
As before, the estimates of the other variables are available from the authors.
The correlation between physical activity rates and obesity is clear; for every
percentage point increase in physical activity rate, the expected obesity rate
drops by $0.43$ percentage points in all four accessiblity specifications.
The indirect effect of physical activity rates on obesity in neighboring zones
is inconsequential, supporting our methodological assertion that public health
variables are not spatially dependent, though the factors leading to them may be.

@@

```{r obesity-models, results="asis"}
texreg(obesity_models[c(2:5)], digits = 4, ci.force = TRUE, 
       booktabs = TRUE, use.packages = FALSE, 
       custom.model.names = c("Size and Distance", "Twitter", "Amenities", "10-Minute Walk"),
       custom.coef.map = list(
         "physact" = "Physical Activity", 
         "lag.physact" = "$\\gamma$: Physical Activity",
         "lambda" = "$\\lambda$: spatial correlation", 
         "access_ls" = "Size and Distance", 
         "multi_ls" = "Amenities", 
         "tweets_ls" = "Twitter Activity", 
         "walk_10TRUE" = "10-minute walk"),
       caption = "Estimated Effect of Accessibility on Physical Activity Rates",
       label = "tab:obesity-models",
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```

@@

## Acknowledgements
This project was funded by the Speedwell Foundation. We are grateful to Rachel
Samuels for helping to assemble the geolocated Twitter data.

>>>>

output: 
  rticles::elsevier_article:
    includes: 
      in_header: preamble.tex
      after_body: appendix.md
csl: elsevier-harvard.csl
bibliography: library.bib
---

@@

The large majority of extant research on the topic of park access and public health
outcomes has been involved with comparing metropolitan regions against each other. For example,
@West2012 used park data from
the Trust for Public Land’s 2010 "City Park Facts" and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine the relationship
between the density of park land and physical activity and obesity rates for 67
metropolitan statistical areas in the U.S. The findings in this case conformed to 
expectation; that is, the study found a significant positive association
between park density and physical activity and a significant negative
association between park density and obesity.
@Larson2016 similarly used self-reported scores on the Gallup-Healthways
Wellbeing Index to evaluate the relationship between physical health and park
quantity, quality, and accessibility in 44 U.S. cities. While the authors found
positive associations between wellbeing and park quality and accessibility,
these relationships were not statistically significant.
Conversely, @Richardson2012 examined the relationship between total urban
greenspace and mortality rates for selected maladies; though the authors did not
find a statistically significant relationship between the quantity of urban
greenspace and mortality caused by lung cancer, diabetes, heart disease or car
accidents, they did find that all-cause mortality was --- oddly --- *higher* in greener
cities. Finally, in a meta-analysis of 20 peer reviewed journal articles exploring the
relationship between parks and objectively measured physical activity,
@Bancroft2015 found that five studies reported a significant positive
association between the two, six studies produced mixed results, and nine
studies found no association at all.

@@

Variations of this isochrone-based framework are easily derived and relatively
common. 
ParkScore [@parkscore2019] is a sophisticated application of this approach where $D$ is a
carefully calculated 10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. In other applications $D$ might be more
variably defined, such as the presence or percent of green space within a
political or statistical boundary [@Mitchell2008; @Stark2014] or within a specific distance
threshold [@Kaczynski2014]. It is also possible to attenuate an isochrone with 
various factors; for instance, @Dias2019 considered road safety perception as an
extenuating factor in the effective distance to a park.

@@

In spite of its flexibility and basis in choice theory, utility-based
accessibility measures have not received much application in the accessibility
literature compared with distance-based or even gravity-based measures.
@Vale2016 present a descriptive classification of active accessibility techniques, and
explicitly dismiss utility-based metrics for incorporating randomness. The
accessibility formula presented in Equation \ref{eq:utility} is the expectation of a
random utility process, and is not in and of itself random any more than the
gravity model is random.
Utility-based accessibility metrics are commonly used, however, in alternatives
analyses of transit infrastructure improvements [@DeJong2007]. A reason for this
is likely that a regional travel demand model is available to the analysts, thus
making calibrated and multimodal logsums readily available [@Geurs2010].

@@

To this dataset of parks and regularized green spaces, we add a polygons layer of
cemeteries in New York that are open to the public. Cemeteries are important
green spaces that can be used for many types of physical activity. These
operations leave us with `r nrow(parks)` distinct green spaces; for simplicity 
we will refer to both parks and cemeteries as "parks" going forward. 
For each park, 
we calculate the size of the park in acres. Finally, using a Twitter application
programming interface (API) implemented with the Python package Tweepy,  we also
collected and stored tweets containing geotags of precise geographic coordinates
located within the boundaries of the parks; from this information,
we were able to segregate tweets that were generated within a park in September
2014. This geolocated twitter activity could provide an additional point of information
on the degree to which individual parks are actually used [@Wang2016].

@@

We calculate a set of utility-based accessibility statistics for each tract
in New York City. The most basic utility specification includes only the
the park size in acres and the distance from the population-weighted tract centroid
to the boundary of the park in miles,
\begin{equation}\label{eq:u_access}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij})
\end{equation}
The logarithmic transform allows for diminishing marginal utility of distance
and park size: A 1-mile increase to a trip matters more for a 1-mile trip than a
10-mile trip. @Macfarlane2019 estimated destination choice 
parameters for park trips in the Alameda County (Oakland), California by applying
this utility specification to passively collected mobile device data, obtaining values of $\lambda_s = 0.373$ and 
$\lambda_d = -1.76$. The ratio of these estimates implies people are willing to
travel roughly six times further to reach a park twice as large. We do not have
access to the necessary data to repeat the @Macfarlane2019 methodology in New
York is cost-prohibitive, and the public parks datasets in Alameda County are
not sufficiently detailed to perform the present accessibility analysis there.
In the absence of park trip destination choice coefficients in New York City, we
adopt these previous estimates.

@Kinnell2006 surveyed park users in New Jersey and applied a multinomial logit
model to determine which factors influence a park's perceived utility as a park
trip destination. By transferring estimates of common variables between our
data and this survey-based model, we can create a second utility specification
as
\begin{equation}\label{eq:u_multi}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \lambda_t * trails + \\
  \lambda_{c} * courts + \lambda_p * playgrounds
\end{equation}
with $\lambda_t = 0.99$, $\lambda_c = 0.43$, and $\lambda_p = 0.26$. Kinnell et al.\
did not transform their size and distance estimates, so we retain the previously 
estimated and applied values. We assume that the other covariates in the 
model which we do not have available to us (e.g., boat launches, picnic areas) are
orthogonal to the included parameters and leaving them out will not affect the
values of the included coefficients.

The third utility specification includes the number of geolocated tweets
emanating from within the park in September of 2014 in addition to the
amenities in Equation \ref{eq:u_multi},
\begin{equation}\label{eq:u_tweets}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \lambda_t * trails + \\
  \lambda_{c} * courts + \lambda_p * playgrounds + \lambda_{tw} * YJ(tweets)
\end{equation}
with $YJ(x)$ being a Yeo-Johnson transformation that implements a diminishing
marginal return similar to $log(x)$, but where $YJ(0) = 0$ [@Yeo2000]. In this
case we have no external information describing how destination choice to
parks may be affected by the twitter activity in the park, so we assert a 
value of $\lambda_{tw} = 0.1$

@@

As the observations are related to each other spatially, it is likely that this
process involves spatial spillovers of one kind or another. A complete treatment of
spatial econometrics is not warranted here; the reader is directed to
@LeSage2009 as well as @LeSage2014.
Suffice it to say that the presence of spatial data generating processes can
negatively affect econometric interpretation in a variety of ways. Each of these
processes rely on spatial autoregression, where elements of some random
variable $x$ are spatially dependent on other elements, 
\begin{equation}\label{eq:autoregression}
 x = \rho W x + \varepsilon
\end{equation}
with the strength of dependence an estimated parameter $\rho$, the structure of 
spatial dependence governed by the asserted matrix $W$ (details of $W$ follow below),
and some random uncorrelated residual $\varepsilon$. A collection of models is
available to represent spatial autodependence in the model residuals, the 
independent variables, and the dependent variable.

@@

Table \ref{tab:pa-models} present the estimated spatial lag and accessibility
coefficients for three models using different specifications of the utility
equation, as well as the ten-minute walk buffer calculated by ParkScore. 
Note that the model called "Size and Distance" is precisely the model
labeled "SDEM" in Table \ref{tab:slx-sdem}. The controlling covariates are
suppressed for clarity but do not change meaningfully across the four specifications.
Complete model estimates are available in the Appendix. Each of the three
utility-based specifications is significant at the $\alpha = 0.05$ level or lower,
and is of the hypothesized directionality. That is, residents living in tracts 
with increased utility-based accessiblity to green spaces show a significantly
higher physical activity participation rate. Taking the rough mean of the three
estimated utility-based coefficient, moving from the least-accessible tract ($A_i = -3$) to the
most-accessible tract ($A_i = 3$) is expected to raise the physical activity rate
by `r 6 * .2` percentage points. With the understanding of the full 95\% confidence
interval, we could suggest that the true effect of excellent versus poor park access
is anywhere between `r 6 * .05` and `r 6 * .35` percentage points.

@@

We now consider the impacts of a model where the dependent variable is
the *obesity* rate in a tract, and physical activity becomes an independent covariate
alongside the controlling variables and the accessibility metrics.
Table \ref{tab:obesity-models} presents estimates of the effect of all three
utility-based accessibility measures as well as the 10-minute walk buffer. 
As before, the estimates of the other variables are available in the Appendix.
The correlation between physical activity rates and obesity is clear; for every
percentage point increase in physical activity rate, the expected obesity rate
drops by $0.43$ percentage points in all four accessiblity specifications.
The indirect effect of physical activity rates on obesity in neighboring zones
is inconsequential, supporting our methodological assertion that public health
variables are not spatially dependent, though the factors leading to them may be.

@@

```{r obesity-models, results="asis"}
texreg(obesity_models[c(2:5)], digits = 4, ci.force = TRUE, 
       booktabs = TRUE, use.packages = FALSE, 
       custom.model.names = c("Size and Distance", "Twitter", "Amenities", "10-Minute Walk"),
       custom.coef.map = list(
         "physact" = "Physical Activity", 
         "lag.physact" = "$\\gamma$: Physical Activity",
         "lambda" = "$\\lambda$: spatial correlation", 
         "access_ls" = "Size and Distance", 
         "multi_ls" = "Amenities", 
         "tweets_ls" = "Twitter Activity", 
         "walk_10TRUE" = "10-minute walk"),
       caption = "Estimated Effect of Accessibility on Obesity Rates",
       label = "tab:obesity-models",
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```

@@

## Acknowledgements
This project was funded by the Speedwell Foundation. We are grateful to Rachel
Samuels and Neda Madi for helping to assemble the geolocated Twitter data.
